Energy Dependence of Synchrotron X-Ray Rims in Tycho's Supernova Remnant

### 2014-12-17 manuscript/paper-tycho.tex
adaptive rminarc bug fix; eta2=1 loss-lim fits
Update paper table w/eta2=1 fit #s (extremely minor changes)
Clean up notes/Qs/etc, minor updates to paper

% TODO define loss-limited properly...
% ~210 words -- a little too long
\begin{abstract}
Several young supernova remnants exhibit thin X-ray bright rims of synchrotron
radiation at their forward shocks.  Thin rims have been taken to indicate that
shock-accelerated electrons rapidly cool downstream of the shock, requiring
strong magnetic field amplification.  But, magnetic field damping immediately
behind the shock could produce similarly thin rims.  We measured rim widths
around Tycho's supernova remnant in 5 energy bands using an archival $750
\unit{ks}$ \Chandra observation.  Rims consistently narrow with increasing
energy and are well-described by either loss-limited or magnetically damped
scenarios, hence X-ray rim widths alone cannot discriminate between models.
Loss-limited fits require amplified magnetic fields of $\abt200$--$1000 \muG$
and particle diffusion coefficients $\abt1$--$100\times$ Bohm values, whereas
magnetic damping permits magnetic fields of $\abt 10$--$300 \muG$ and
comparatively smaller diffusion coefficients (typically $\abt 1$--$10$).
Corresponding thin rim radio rims require magnetic damping, however.  We
jointly model radio and X-ray profiles and estimate magnetic damping length of
order $1$--$5$\% of remnant radius, with magnetic field strengths $\abt
50$--$400 \muG$ assuming Bohm diffusion.  Diffusion coefficients, magnetic
field strengths, and magnetic damping lengthscales are somewhat degenerate in
setting observed rim widths and drop-offs.
Although the energy dependence of synchrotron rims cannot discriminate magnetic
field structure, joint radio and X-ray modeling favors magnetic damping in
Tycho's SNR.
\end{abstract}

@@

% NOTE go through and double check tenses.
We measured synchrotron rim full widths at half maximum (FWHMs) from an
archival \Chandra ACIS-I observation of Tycho
(RA: 00\tsup{h}25\tsup{m}19\fs0, dec: +64\arcdeg08\arcmin10\farcs0; J2000)
between 2009 Apr 11 and 2009 May 5 (PI: J. Hughes;
\dataset[ADS/Sa.CXO\#obs/10093--10097]{ObsIDs: 10093--10097},
\dataset[ADS/Sa.CXO\#obs/10902--10906]{10902--10906}).
The total exposure time was $734 \unit{ks}$.
Level 1 \Chandra data were reprocessed with CIAO 4.6 and CALDB 4.6.1.1 and kept
unbinned with ACIS spatial resolution $0.492\arcsec$.
Merged and corrected events were divided into five energy bands:
0.7--1 keV, 1--1.7 keV, 2--3 keV, 3--4.5 keV, and 4.5--7 keV.
We excluded the 1.7--2 keV energy range to avoid \ion{Si}{13} (He$\alpha$)
emission at 1.85 keV, prevalent in the remnant's thermal ejecta, which
might contaminate our nonthermal profile measurements.

@@

% Best fit model results and tables for Regions 1, 16.
% Explain a bit about model shapes, and the relative similarity in shapes
Modeled profiles with best fit FWHMs for Regions 1 and 16 are plotted in
Figure~\ref{fig:rims}, and their derived best-fit FWHM-energy curves
are plotted in Figure~\ref{fig:fits} alongside measured FWHMs.
The best fit model profiles do not quite match our measured rim profiles
(Figure~\ref{fig:profiles}).  Our transport equation does not account for
diffusive transport upstream of the shock, projection effects from shock
ripples or other structure, and thermal emission downstream of the thin rim.
Hence, fitting to FWHM measurements should be less sensitive to these effects
as compared to directly fitting entire rim profiles to model predictions.
We also note that although Figure~\ref{fig:rims} plots model profiles for
$\mu = 1$ only, the best-fit FWHM-energy curves show that model profiles for
$\mu \neq 1$ give very similar FWHM-energy dependence.  This is borne out in
our best fit parameters in Tables~\ref{tab:fits-loss} and \ref{tab:fits-damp}.

@@

\subsubsection{Effect of Tycho distance estimates on model fits}
% TODO need to update for damping results!!!!!

In our modeling we adopted a distance $d$ to Tycho of $3 \unit{kpc}$, but
various estimates for Tycho's distance range between $2.3$--$4 \unit{kpc}$
\citep{hayato2010}.  A larger remnant distance would increase both physical
filament widths and shock velocity estimates from proper motion (as used in our
models).

% lad/ldiff scalings give eta/B0 scalings
In the loss-limited model, we find roughly that $\eta_2 \propto d^2$ with
$B_0$ approximately constant.  The advective lengthscale,
equation~\eqref{eq:lad}, may be rearranged to obtain:
\begin{align}
    B_0 =\; &(317 \muG) \left(\frac{v_d}{10^8 \unit{cm/s}}\right)^{2/3}
                \nonumber \\
            &\times \left(\frac{l_{\mt{ad}}}{0.01 \unit{kpc}}\right)^{-2/3}
                \left(\frac{h\nu}{1 \unit{keV}}\right)^{-1/3}
\end{align}
% TODO if we keep this equation, double check the 317 \muG carefully...
or, more simply,
\begin{equation}
    B_0 \propto \left(v_d\right)^{2/3}
                \left(l_{\mt{ad}}\right)^{-2/3} \nu^{-1/3} .
\end{equation}
Both $l_{\mt{ad}}$ and $v_d$ scale linearly with remnant distance $d$ and thus
their effects cancel in determining the magnetic field.  If diffusion is the
primary control on filament lengthscales, equation~\eqref{eq:ldiff} yields:
\begin{equation}
    \eta \propto \left(l_{\mt{diff}}\right)^2 B_0^{3} \nu^{-(\mu - 1)/2}
\end{equation}
Recall that $\eta \propto \eta_2$ since $E_h = E_2$ is constant for a given
$B_0$, from equation~\eqref{eq:diffcoeff}.  Similar results, primarily for
magnetic field $B_0$, were previously given by \citet{parizot2006}.

In practice, both advection and diffusion are important to our filament widths,
but both scalings $\eta_2 \propto d^2$ and $B_0$ constant occur in model fits
with varying distance.  When comparing full model fits with remnant distances
of $3 \unit{kpc}$ and $4 \unit{kpc}$, the deviation from the idealized scaling
is $\lesssim 1 \%$ for $B_0$ and $\abt 1$--$5\%$ for $\eta_2$.  As varying
remnant distance $d$ leaves width-energy scaling $\mE$ invariant, the relative
contributions of $l_{\mt{ad}}$ and $l_{\mt{diff}}$ should also be invariant.
Then both lengthscales should scale simultaneously with $d$, yielding the
observed behavior.

% Damping and remnant distance
In the damped model, a larger distance $d$ will require larger damping lengths
from magnetic turbulence.  But, fitted values of $a_b$ will remain unchanged as
we report $a_b$ in units of shock radius $r_s$.

@@

Nevertheless, our conclusion that thin radio rims require magnetic damping
should be robust and independent of model assumptions.
\citet{cassam-chenai2007}, for example, observed the same dichotomy in radio
emission for constant and damped magnetic fields in modeled profiles of Tycho's
shock from a self-similar hydrodynamic model with diffusive shock acceleration.
Our model captures the essential physics for a very young remnant such as
Tycho, for which the forward shock should be adiabatic.

@@

The exponential variation in $D(x)$ (equation~\eqref{eq:ddamp}) may not be
physically reasonable; as we have noted above, the assumption $D(x)
\propto 1 / B^2(x)$ gives rise to sharp gradients in diffusion coefficient and
even inverts the effect of $\eta_2$ on our model profiles (where larger
$\eta_2$ can cause rim widths to narrow).  Nevertheless, the modeled behavior
appears physically reasonable.  And, only X-ray profiles are impacted by the
assumption on $D(x)$ as radio profiles assume no diffusion.

% TODO question: we have been ignoring intensity.  What if, e.g.,
% with strong diffusion particles get out of the damping zone faster and don't
% even have a chance to radiate, or something like that?
% I have no clue... check this, double check Sean's normalizations to be safe

>>>>

\begin{abstract}
Several young supernova remnants exhibit thin X-ray bright rims of synchrotron
radiation at their forward shocks.  Thin rims require strong magnetic field
amplification if rim widths are limited by shock-accelerated electron energy
losses.  But, magnetic field damping immediately behind the shock could produce
similarly thin rims.  We measured rim widths around Tycho's supernova remnant
in 5 energy bands using an archival $750 \unit{ks}$ \Chandra observation.  Rims
consistently narrow with increasing energy and are well-described by either
loss-limited or magnetically damped scenarios, hence X-ray rim widths alone
cannot discriminate between models.  Loss-limited rims require amplified
magnetic fields of $\abt200$--$1000 \muG$ and particle diffusion coefficients
$\abt1$--$100\times$ Bohm values, whereas magnetic damping permits magnetic
fields of $\abt 10$--$300 \muG$ and comparatively smaller diffusion
coefficients (typically $\abt 1$--$10$).  Corresponding thin rim radio rims
require magnetic damping, however.  We jointly model radio and X-ray profiles
and estimate magnetic damping length of order $1$--$5$\% of remnant radius,
with magnetic field strengths $\abt 50$--$400 \muG$ assuming Bohm diffusion.
Although the energy dependence of synchrotron rims cannot discriminate magnetic
field structure, joint radio and X-ray modeling favors magnetic damping in
Tycho's SNR.
\end{abstract}

@@

We measured synchrotron rim full widths at half maximum (FWHMs) from an
archival \Chandra ACIS-I observation of Tycho
(RA: 00\tsup{h}25\tsup{m}19\fs0, dec: +64\arcdeg08\arcmin10\farcs0; J2000)
between 2009 Apr 11 and 2009 May 5 (PI: J. Hughes;
\dataset[ADS/Sa.CXO\#obs/10093--10097]{ObsIDs: 10093--10097},
\dataset[ADS/Sa.CXO\#obs/10902--10906]{10902--10906}).
The total exposure time was $734 \unit{ks}$.
Level 1 \Chandra data were reprocessed with CIAO 4.6 and CALDB 4.6.1.1 and kept
unbinned with ACIS spatial resolution $0.492\arcsec$.
Merged and corrected events were divided into five energy bands:
0.7--1 keV, 1--1.7 keV, 2--3 keV, 3--4.5 keV, and 4.5--7 keV.
We excluded the 1.7--2 keV energy range to avoid \ion{Si}{13} (He$\alpha$)
emission at 1.85 keV, prevalent in the remnant's thermal ejecta, which
might contaminate our nonthermal profile measurements.

@@

% Best fit model results and tables for Regions 1, 16.
% Explain a bit about model shapes, and the relative similarity in shapes
Modeled profiles with best fit FWHMs for Regions 1 and 16 are plotted in
Figure~\ref{fig:rims}, and their derived best-fit FWHM-energy curves
are plotted in Figure~\ref{fig:fits} alongside measured FWHMs.
The best fit model profiles do not quite match measured rim profiles
(Figure~\ref{fig:profiles}).  Our transport equation does not account for
diffusive transport upstream of the shock, projection effects from shock
ripples or other structure, and thermal emission downstream of the thin rim.
Fitting to multiple FWHM measurements should be robust to these effects as
compared to directly fitting entire rim profiles to model predictions.
Although Figure~\ref{fig:rims} plots model profiles for $\mu = 1$ only, the
best-fit FWHM-energy curves show that model profiles for $\mu \neq 1$ give very
similar FWHM-energy dependence.

@@

\subsubsection{Effect of Tycho distance estimates on model fits}

Our models adopted a distance $d$ to Tycho of $3 \unit{kpc}$, but estimates for
Tycho's distance range between $2.3$--$4 \unit{kpc}$ \citep{hayato2010}.  A
larger remnant distance would increase both physical filament widths and shock
velocity estimates from proper motion; we may then derive expected scalings for
modeled $\eta_2$ and $B_0$ as functions of assumed distance $d$.  The advective
lengthscale, equation~\eqref{eq:lad}, may be rearranged to obtain:
\begin{align}
    B_0 =\; &(317 \muG) \left(\frac{v_d}{10^8 \unit{cm/s}}\right)^{2/3}
                \nonumber \\
            &\times \left(\frac{l_{\mt{ad}}}{0.01 \unit{kpc}}\right)^{-2/3}
                \left(\frac{h\nu}{1 \unit{keV}}\right)^{-1/3}
\end{align}
% TODO if we keep this equation, double check the 317 \muG carefully...
or, more simply,
\begin{equation}
    B_0 \propto \left(v_d\right)^{2/3}
                \left(l_{\mt{ad}}\right)^{-2/3} \nu^{-1/3} .
\end{equation}
Both $l_{\mt{ad}}$ and $v_d$ scale linearly with remnant distance $d$ and thus
their effects cancel in determining the magnetic field.  If diffusion is the
primary control on filament lengthscales, equation~\eqref{eq:ldiff} yields:
\begin{equation}
    \eta \propto \left(l_{\mt{diff}}\right)^2 B_0^{3} \nu^{-(\mu - 1)/2}
\end{equation}
Recall that $\eta \propto \eta_2$ since $E_h = E_2$ is constant for a given
$B_0$, from equation~\eqref{eq:diffcoeff}.  Similar results, primarily for
magnetic field $B_0$, were previously given by \citet{parizot2006}.

In practice, model fits with varying distance obey both scalings $\eta_2
\propto d^2$ and $B_0$ constant.  When comparing full model fits with remnant
distances of $3 \unit{kpc}$ and $4 \unit{kpc}$, the deviation from the
idealized scaling is $\lesssim 1 \%$ for $B_0$ and $\abt 1$--$5\%$ for
$\eta_2$.  As varying remnant distance $d$ leaves width-energy scaling $\mE$
invariant, the relative contributions of $l_{\mt{ad}}$ and $l_{\mt{diff}}$
should also be invariant.  Then both lengthscales should scale simultaneously
with $d$, yielding the observed behavior.  In the damped model, a larger
distance $d$ will require larger physical damping lengths from magnetic
turbulence.  But, fitted values of $a_b$ should remain unchanged as we report
$a_b$ in units of shock radius $r_s$.

@@

Nevertheless, our conclusion that thin radio rims require magnetic damping is
supported by more sophisticated modeling.  Hydrodynamic models with diffusive
shock acceleration \citep{cassam-chenai2007, slane2014} also cannot produce
radio profiles with narrow rims in a purely advected magnetic field.
\citet{cassam-chenai2007} produced radio rims using a damped magnetic field,
though they could not simultaneously reconcile predicted X-ray and radio
intensities to observations.  Our model captures the essential physics for a
very young remnant such as Tycho, for which the forward shock should be
adiabatic.

@@

The exponential variation in $D(x)$ (equation~\eqref{eq:ddamp}) may not be
physically reasonable; as noted above, the assumption $D(x)
\propto 1 / B^2(x)$ gives rise to sharp gradients in diffusion coefficient and
even inverts the effect of $\eta_2$ on our model profiles (where larger
$\eta_2$ can cause rim widths to narrow).  Nevertheless, the modeled behavior
appears physically reasonable.  And, only X-ray profiles are impacted by the
assumption on $D(x)$ as radio profiles assume no diffusion.

% TODO question: we have been ignoring intensity (absolute comparison).  What
% if, e.g., with strong diffusion particles get out of the damping zone faster
% and don't even have a chance to radiate?
% I have no clue... check this, double check Sean's normalizations to be safe
